---
title: "D2 Data Transformation"
author: "William Stingl"
date: "9/21/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 5.1 Introduction

##### 5.1.1 Prerequisites

```{r load_libraries}
library(nycflights13)
library(tidyverse)
```

##### 5.1.2 nycflights13

```{r inspect_flights}
flights
```

##### 5.1.3 dplyr basics

In this chapter you are going to learn the five key dplyr functions that allow you to solve the vast majority of your data manipulation challenges:

*Pick observations by their values (filter()).
*Reorder the rows (arrange()).
*Pick variables by their names (select()).
*Create new variables with functions of existing variables (mutate()).
*Collapse many values down to a single summary (summarize()).

These can all be used in conjunction with group_by() which changes the scope of each function from operating on the entire dataset to operating on it group-by-group. These six functions provide the verbs for a language of data manipulation.

All verbs work similarly:

* 1. The first argument is a data frame.
* 2. The subsequent arguments describe what to do with the data frame, using the variable names (without quotes).
* 3. The result is a new data frame.

Together these properties make it easy to chain together multiple simple steps to achieve a complex result. Let’s dive in and see how these verbs work.

### 5.2 Filter rows with filter()

filter() allows you to subset observations based on their values. The first argument is the name of the data frame. The second and subsequent arguments are the expressions that filter the data frame. For example, we can select all flights on January 1st with:

```{r example_1}
filter(flights, month == 1, day == 1)
```

```{r example_2}
jan1 <- filter(flights, month == 1, day == 1)
```

```{r example_3}
jan1
```

```{r example_4}
(jan1 <- filter(flights, month == 1, day == 1))
```

##### 5.2.2 Comparisons

To use filtering effectively, you have to know how to select the observations that you want using the comparison operators. R provides the standard suite: >, >=, <, <=, != (not equal), and == (equal).

When you’re starting out with R, the easiest mistake to make is to use = instead of == when testing for equality.

```{r example_5}
#filter(flights, month = 1)
```

There’s another common problem you might encounter when using ==: floating point numbers. These results might surprise you!

```{r example_6}
sqrt(2) ^ 2 == 2
```

Computers use finite precision arithmetic (they obviously can’t store an infinite number of digits!) so remember that every number you see is an approximation. Instead of relying on ==, use near():

##### 5.2.2 Logical operators

Multiple arguments to filter() are combined with “and”: every expression must be true in order for a row to be included in the output. For other types of combinations, you’ll need to use Boolean operators yourself: & is “and”, | is “or”, and ! is “not”.

The following code finds all flights that departed in November or December:

```{r example_7}
filter(flights, month == 11 | month == 12)
```

The order of operations doesn’t work like English. You can’t write filter(flights, month == (11 | 12)), which you might literally translate into “finds all flights that departed in November or December”. Instead it finds all months that equal 11 | 12, an expression that evaluates to TRUE. In a numeric context (like here), TRUE becomes one, so this finds all flights in January, not November or December. This is quite confusing!

A useful short-hand for this problem is x %in% y. This will select every row where x is one of the values in y. We could use it to rewrite the code above:

```{r example_8}
(nov_dec <- filter(flights, month %in% c(11, 12)))
```

Sometimes you can simplify complicated subsetting by remembering De Morgan’s law: !(x & y) is the same as !x | !y, and !(x | y) is the same as !x & !y. For example, if you wanted to find flights that weren’t delayed (on arrival or departure) by more than two hours, you could use either of the following four filters:

```{r example_9}
filter(flights, !(dep_delay > 120) & !(arr_delay > 120))
filter(flights, !(dep_delay > 120), !(arr_delay > 120)) #same as above, but with commas
filter(flights, !(arr_delay > 120 | dep_delay > 120)) #De Morgan's Law
filter(flights, arr_delay <= 120 & dep_delay <= 120)
```

##### 5.2.3 Missing values

One important feature of R that can make comparison tricky are missing values, or NAs (“not availables”). NA represents an unknown value so missing values are “contagious”: almost any operation involving an unknown value will also be unknown.

```{r example_10}
100 == NA
2 > NA
x <- NA
is.na(x)
```

How do I get rid of NA's on my dataset?

```{r example_11}
filter(flights, is.na(dep_time))
filter(flights, !is.na(dep_time))

```

##### 5.2.4 Exercises

> 1. Find all flights that

Had an arrival delay of two or more hours

```{r exercise_5.2.1.a}
filter(flights, arr_delay >= 120)
```

Flew to Houston (IAH or HOU)

```{r exercise_5.2.1.b}
filter(flights, dest == 'IAH' | dest == 'HOU')
filter(flights, dest %in% c('IAH', 'HOU'))
```

Were operated by United, American, or Delta

```{r exercise_5.2.1.c}
filter(flights, carrier == 'UA' | carrier == 'AA' | carrier == 'DL')
filter(flights, carrier %in% c('UA', 'AA', 'DL'))
```

Departed in summer (July, August, and September)

```{r exercise_5.2.1.d}
filter(flights, month %in% c(7, 8, 9))
# or
filter(flights, month %in% 7:9)
```

Arrived more than two hours late, but didn’t leave late

```{r exercise_5.2.1.e}
filter(flights, arr_delay > 120 & dep_delay <= 0)
```

Were delayed by at least an hour, but made up over 30 minutes in flight

```{r exercise_5.2.1.f}
filter(flights, dep_delay >= 60 & dep_delay - arr_delay > 30) # not quite sure this is correct
filter(flights, dep_delay >= 60 & arr_delay < -30)
```

Departed between midnight and 6am (inclusive)

```{r exercise_5.2.1.g}
filter(flights, between(dep_time, 00, 600))
```

> 2. Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?

The expression between(x, left, right) is equivalent to x >= left & x <= right.

Of the answers in the previous question, we could simplify the statement of departed in summer (month >= 7 & month <= 9) using the between() function.

```{r exercise_5.2.2}
filter(flights, between(month, 7, 9))
```

> 3. How many flights have a missing dep_time? What other variables are missing? What might these rows represent?

```{r exercise_5.2.3.a}
filter(flights, is.na(dep_time))
```

```{r exercise_5.2.3.b}
summary(flights)
```


### 3.4 Select Columns with select()

```{r}
select(flights, year, month, day)
select(flights, year:day)

select(flights, -(year:day))
select(flights, dep_time:time_hour)
```

```{r}
select(flights, starts_with('dep'))
select(flights, ends_with('delay'))
select(flights, contains('_'))
```

```{r}
rename(flights, tail_num = tailnum)
```

```{r}
select(flights, dep_delay, arr_delay, everything())
```



### 5.5 Add new variable with mutate()

creates new column

```{r}
flights_narrow <- select(flights, year:day, ends_with('delay'), distance, air_time)

flights_narrow
```


```{r}
mutate(flights_narrow, gain = dep_delay - arr_delay, speed = distance/air_time * 60)
```


```{r}
mutate(flights_narrow, gain = dep_delay - arr_delay, speed = distance/air_time * 60, gain_per_hour = gain/ air_time)
```

```{r}
transmute(flights, gain = dep_delay - arr_delay, speed = distance/ air_time * 60, gain_per_hour = gain/ air_time)
```


### 5.6 Grouped summaries with summarise()

It collapses a data frame to a single row

```{r}
summarize(flights, delay = mean(dep_delay, na.rm = TRUE))
```

```{r}
by_day <- group_by(flights, year, month, day)

summarise(by_day, delay = mean(dep_delay, na.rm = TRUE))
``` 

```{r}
by_carrier <- group_by(flights, carrier)

summarize(by_carrier, delay = mean(dep_delay, na.rm = TRUE))
```

```{r}
by_airTime <- group_by(flights, air_time)

summarize(by_airTime, delay = mean(dep_delay, na.rm = TRUE))
```

```{r}
by_airport <- group_by(flights, origin)

summarize(by_airport, delay = mean(dep_delay, na.rm=TRUE))
```

### 5.6.1 Combinign multiple operations with the pipe
Imagine that we want to explore the relationship between the distance and average delay for each location, using what you know about dplyr.

groups flight by destination

```{r}
by_dest <- group_by(flights, dest)
```

summarize the group using average distance, average delay, an count:

```{r}
delays <- summarise(by_dest,
      count = n(),
      dist = mean(distance, na.rm=TRUE),
      delay = mean(arr_delay, na.rm=TRUE))

delays
```

```{r}
ggplot(data = delays) + 
  geom_point(mapping = aes(x = dist, y = delay, size = count))
```

```{r}
delays <- filter(delays, count > 20, dest != "HNL")
delays
```

```{r}
ggplot(data = delays) + 
  geom_point(mapping = aes(x = dist, y = delay, size = count))
```


```{r}
ggplot(data = delays, mapping = aes(x = dist, y = delay, size = count)) +
  geom_point(mapping = aes(size = count), alpha = 1/3) +
  geom_smooth(se = FALSE)
```

```{r}
delays <- flights %>% 
  group_by(dest) %>%
  summarize(
    count = n(),
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE)
  ) %>% 
  filter(count > 20, dest != "HNL")

ggplot(data = delays, mapping = aes(x = dist, y = delay, size = count)) +
  geom_point(mapping = aes(size = count), alpha = 1/3) +
  geom_smooth(se = FALSE)
```

###### 5. 6. 2 Missing values

```{r}
flights %>%
  group_by(year, month, day) %>%
  summarize(delay = mean(dep_delay))
```

```{r}
flights %>%
  group_by(year, month, day) %>%
  summarize(delay = mean(dep_delay, na.rm=TRUE))
```


















