---
title: "DataVisualisation"
author: "William Stingl"
date: "8/21/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r cars}
summary(cars)
```


# Data Visualization

### 3.1 Introduction

```{r}
# load tidyverse library everyday

library(tidyverse)
```

### 3.2 First Steps 

QUESTIONS: Do cars with big engines use more gas than cars with small engines?

###### 3.2.1 The mpg data frame

 A data frame is a rectangular collection of variables (in the columns) and observations (in the rows). mpg contains observations collected by the US Environmental Protection Agency on 38 models of car.

```{r}
mpg

```

### 3.2.2 Creating a ggplot

```{r}
ggplot(data=mpg) + geom_point(mapping = aes(x = displ, y = hwy))
```

##### Excercises

1. Run ggplot(data = mpg). What do you see?

```{r}
ggplot(data=mpg)
```

2. How many rows are in mpg? How many columns?
11
```{r}
?mpg
```

3. What does the drv variable describe? Read the help for ?mpg to find out.
drive
```{r}

```

4. Make a scatterplot of hwy vs cyl.

```{r}
ggplot(data=mpg) + geom_point(mapping = aes(x = cyl, y = hwy))
```

5. What happens if you make a scatterplot of class vs drv? Why is the plot not useful?

```{r}
ggplot(data=mpg) + geom_point(mapping = aes(x = class, y = drv))
```

### 3.3 Aesthetic Mappings

```{r}
ggplot(data=mpg) + geom_point(mapping = aes(x = displ, y = hwy, color=class))
```

```{r}
ggplot(data=mpg)+geom_point(mapping=aes(x=displ, y=hwy, size=class))
```

```{r}
ggplot(data=mpg)+geom_point(mapping=aes(x=displ, y=hwy, alpha=class))
```

```{r}
ggplot(data=mpg)+geom_point(mapping=aes(x=displ,y=hwy, shape=class))
```

```{r}
ggplot(data=mpg)+geom_point(mapping=aes(x=displ, y=hwy),color="blue")
```

###### 3.3.1 EXCERCISES

1. What’s gone wrong with this code? Why are the points not blue?
color="blue" has to be outside the argument
```{r}
ggplot(data=mpg)+geom_point(mapping=aes(x=displ, y=hwy),color="blue")
```

2. Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset). How can you see this information when you run mpg?
The variables that are categorical include characters, while the continuous variable are defined by integers and decimals. 
```{r}
?mpg
```

3. Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?

```{r}
ggplot(data=mpg)+geom_point(mapping=aes(x=displ, y=hwy, color=cty))
```

```{r}
ggplot(data=mpg)+geom_point(mapping=aes(x=displ, y=hwy, size=cty))
```

A continuous variable cannot be mapped to shape.


4. What happens if you map the same variable to multiple aesthetics?
redundant
```{r}
ggplot(data=mpg)+geom_point(mapping=aes(x=displ, y=hwy, color=hwy))
```

5. What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point)

```{r}

```

6. What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)? Note, you’ll also need to specify x and y.

```{r}
ggplot(data=mpg)+geom_point(mapping=aes(x=displ, y=hwy, color=displ<5))
```

### 3.5 Facets
Inside of the sesthetic function the argument is about identifying the variables then another other that can describe them. 
```{r}
ggplot(data=mpg)+geom_point(mapping=aes(x=displ, y=hwy))+facet_wrap(~class, nrow=2)
```
Facets are for categorical variables
```{r}
ggplot(data=mpg)+geom_point(mapping=aes(x=displ, y=hwy))+facet_grid(drv~cyl)
```

### 3.5.1 Excercises

1. What happens if you facet on a continuous variable?
Data becomes very murky and hard to understand. Without categories, adding another conitnuous varible muddles how useful the graph becomes. Good for identifying trends.
```{r}
ggplot(data=mpg)+geom_point(mapping=aes(x=displ, y=hwy))+facet_wrap(~cty, nrow=1)
```

2. What do the empty cells in plot with facet_grid(drv ~ cyl) mean? How do they relate to this plot?
They represent the certain cars that are really produced. For example, 4 and 5-cylinder-rear wheel drive cars and 5 cylinder 4-wheel-drive cars are rare if not non existent.
```{r}
ggplot(data=mpg)+geom_point(mapping=aes(x=displ, y=hwy))+facet_grid(drv~cyl)
```

3. What plots does the following code make? What does . do?
When the decimal is on the right side of the facet_grid argument the displ,hwy graph becomes organized by rows of the variable(drv) in the facet_grid function.
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + facet_grid(drv ~ .)
```

When the decimal is on the left side of the facet_grid argument the displ,hwy graph becomes organized by columns of the variable(cyl) in the facet_grid function.
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + facet_grid(. ~ cyl)
```

4.Take the first faceted plot in this section:
Faceting seems like r's way of arranging and specifying more unique data sets. If the graph is paired with a categorical variable the data becomes more clear and specified. Also, creating x amount of rows to model the position of the graphs as well are an advantadge. However, graphs stacked on top of each other lose the x axis counter, and data can become clustered up together into black lumps, which can make distinguishing the data visually pretty imprecise.
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + facet_wrap(~ class, nrow = 2)
```

5. Read ?facet_wrap. What does nrow do? What does ncol do? What other options control the layout of the individual panels? Why doesn’t facet_grid() have nrow and ncol arguments?

nrow- number of rows to organize the data
ncol- number of columns we would like to organize the data by
levels

```{r}
?facet_wrap
```

6. When using facet_grid() you should usually put the variable with more unique levels in the columns. Why?

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + facet_grid(cyl~cty)
```

### 3.6 Geometric Objects

```{r}
ggplot(data=mpg)+geom_smooth(mapping = aes(x=displ, y=hwy))
```

```{r}
ggplot(data = mpg) + geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))
```

```{r}
ggplot(data = mpg) + geom_smooth(mapping = aes(x = displ, y = hwy, group = drv))
```

```{r}
ggplot(data = mpg) + geom_smooth(mapping = aes(x = displ, y = hwy, color = drv))
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + geom_smooth(mapping = aes(x = displ, y = hwy))
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = drv)) + geom_smooth(mapping = aes(x = displ, y = hwy, color = drv))
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + geom_point() + geom_smooth()
```


```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + geom_point(mapping = aes(color = class)) + geom_smooth()
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + geom_point(mapping = aes(color = class)) + geom_smooth(data = filter(mpg, class == "subcompact"), se = FALSE)
```

1. What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?

geom_abline is horizontal, vertical, and diagonal, geom_boxplot, geom_freqpoly, geom_area

```{r}
??geom 
```

2. Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.

se=False takes out the error parameters around the line graph.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + geom_point() + geom_smooth(se = FALSE)
```

3. What does show.legend = FALSE do? What happens if you remove it? Why do you think I used it earlier in the chapter?

Show.legend = FALSE removes the legend on the side of the graph. Focus more on the graph.
```{r}
ggplot(data = mpg) + geom_smooth(mapping = aes(x = displ, y = hwy, color = drv), show.legend = FALSE)
```

4. What does the se argument to geom_smooth() do?
it just adds and removes the error parameters
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + geom_point() + geom_smooth(se = FALSE)
```

5. Will these two graphs look different? Why/why not?
No, because you are telling the program to fill in the empty brackets with the previous or subsequent arguments, which in this case is "data = mpg, mapping = aes(x = displ, y = hwy."
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))

```


### 3.7 Statistical transformation

```{r}
diamonds
```


```{r}
ggplot(data = diamonds) + geom_bar(mapping = aes(x = cut))
```

### 3.8 Position Adjustments

```{r}
ggplot(data = diamonds) + geom_bar(mapping = aes(x = cut, color = cut))
```

```{r}
ggplot(data = diamonds) + geom_bar(mapping = aes(x = cut, fill = cut))
```

```{r}
ggplot(data=diamonds) + geom_bar(mapping = aes(x = cut, fill = clarity))
```

```{r}
ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) + geom_bar(alpha = .2, position = "identity")
```

```{r}
ggplot(data = diamonds, mapping = aes(x = cut, colour = clarity)) + geom_bar(fill = NA, position = "identity")
```

```{r}
ggplot(data = diamonds) + geom_bar(mapping = aes(x = cut, fill = clarity), position = "fill")
```

```{r}
ggplot(data = diamonds) + geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge")
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy), position = "jitter")
```

# 3.8.1 Excercises

Notebook Check  - exercises & notes

Project - start next week due after


```{r}

```








